---
title: "SCMC - Stochastic Collocation Monte Carlo"
output:
  github_document:
    md_extensions: -tex_math_dollars
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a development version of the `scmc` package for the R programming
language.

## Installation

The package isn't available on CRAN so the only way to install the package is to
use the `devtools` package and run

```{r, eval = FALSE}
devtools::install_github("blaza/scmc")
```

## Usage

The main function currently implemented is `univariate_sampler` which is a
flexible implementation of the method (and thus with a bit more complicated
interface) for generating univariate distributions.
We'll cover here a couple of basic examples which give an overall picture of the
package capabilities.

### Example: Logistic distribution

We'll generate variates from the [Logistic distribution](https://en.wikipedia.org/wiki/Logistic_distribution).

The SCMC method implies interpolating $F_Y^{-1}(F_X(x))$, where $Y$ is the
target random variable and $X$ is a random variable which can be efficiently
generated, and generating the samples $y_i,\ i=1,\dots,n$ using the formula
$y_i = F_Y^{-1}(F_X(\xi_i))$, where $\xi_i$ are variates from the $X$ distribution.
In this example we'll use the standard normal variable $X$. By default, we use
the `RcppZiggurat::zrnorm` function to generate normal variates.

The code to generate the Logistic distribution in the `scmc` package is

```{r}
library(scmc)
# create the sampler
sampler <- univariate_sampler(qlogis, gaussian_nodes(7))

# generate 10000 random variates
smp <- sampler(1e5)
```

In its basic form, the `univariate_sampler` function requires the inverse $F_Y^{-1}$ (i.e. the quantile function of $Y$) as the first argument, and the nodes for the interpolation. In cases where normally distributed $X$ are used, optimal nodes for interpolation are the nodes of the Gaussian quadrature with respect to the weight function $f_X(x)$ (density of $X$). The third argument to `univariate_sampler` is `xdist` which is by default `"norm"`, indicating the standard normal distribution.

The quality of the generated sample can be visualized with it's density
```{r}
# plot the sample density
plot(density(smp))

# add a plot of the theoretical logistic density
curve(dlogis(x), add = TRUE, col = "green")
```

The curves are nearly the same, so the approximation is good.

